---
title: "Data Science Workshop Full Guide"
author: "STEM Success Team"
date: "2025-02-20"
output: html_document
---
Don't forget to replace "pathtofile" with your own file path to the dataset on your PC! 
Import data 
```{r}
df <- read.csv("pathtofile", header = TRUE)
```

```{r}
str(df)
```

Check for total missing values and in what columns they reside in
```{r}
sum(is.na(df))

colSums(is.na(df))

nrow(df)
```

Use na.omit to remove rows with missing data
```{r}
noMissingValues_df <- na.omit(df)

nrow(noMissingValues_df)
```

How many duplicates?
```{r}
dupeSum <- sum(duplicated(noMissingValues_df))
dupeSum
```

Remove duplicates
```{r}
noDupes_df <- unique(noMissingValues_df)

rowCount <- nrow(noDupes_df)
rowCount
```

```{r}
print(head(noDupes_df))
```

Binary encoding 
```{r}
binaryEncoded_df <-noDupes_df
binaryEncoded_df$Extracurricular.Activities <- ifelse(binaryEncoded_df$Extracurricular.Activities == "Yes", 1, 0)

clean_df <- binaryEncoded_df
print(head(clean_df))
```

Remove target variable for multicollinearity analysis
```{r}
variable_df <- clean_df[, !names(clean_df) %in% 'Performance.Index']
print(head(variable_df))
```

Create correlation matrix
```{r}
corrmat <- cor(variable_df)
print(corrmat)
```

Creat heatmap for better visualization
```{r}
library(corrplot)
corrplot(corrmat, method = "color", type = "full", 
         col = colorRampPalette(c("red", "gold", "blue"))(200),
         tl.cex = 0.6, tl.col = "black", title = "Correlation Heatmap")
```

Build model
```{r}
studyTime <- clean_df$Hours.Studied
scores <- clean_df$Previous.Scores
extra <- clean_df$Extracurricular.Activities
sleep <- clean_df$Sleep.Hours
practice <- clean_df$Sample.Question.Papers.Practiced
grade <- clean_df$Performance.Index

model <- lm(grade ~ studyTime + scores + extra + sleep + practice, data = clean_df)

summary(model)
```

Model analysis
```{r}
plot(model)
```
Variance Inflation Factors (VIF)
```{r}
library(car)
vif(model)
```






